\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-01 draft.tex
compile draft with updated game-tree
revise text for straightforward presentation of one-sided break

Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country. First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value. Next, the uncertainty about the the median legislator's identity is resolved. This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously. The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.

@@

I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small. This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}

@@

The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
\begin{equation}
    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
  \label{eq:jv2}
\end{equation}
where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$. Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement. One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.

>>>>

Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.

Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement. I assume that these probabilities are less than one-half and mutually exclusive.\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}

After this, the uncertainty about the the median legislator's identity is resolved. This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously. The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.

@@

@@

The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
\begin{equation}
    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
  \label{eq:jv2}
\end{equation}
where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$. Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement.
